---
title: "Initial Korean covid 19"
author: "Leah Castro"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

When I made the graphs, I saw that a lot more people were alive than dead. The percentage chart showed that of the people who died, about two-thirds were men and one-third were women, meaning the disease killed more men. Among those who lived, most were women (about 57%) and fewer were men (about 43%).  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The numerical results back up what we saw in the graphs. About 1.4% of women died from the disease, while 3% of men did which is double the women’s rate. This shows that men have a higher chance of dying than women.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

When looking at the results, I saw that men in South Korea are 2.3 times more likely to die from COVID than women. This matches what we’ve seen before, where most of the COVID deaths were men. I also noticed the P-value is very low (0.0009), which means the chances of men and women dying from COVID are not the same because men have a much higher risk than women.

## Conclusions

